Efficiently stable presentations from error-correcting codes

### 2023-11-01 efficient_stability.tex
cleared notes

\begin{abstract}
\mnote{After reading the abstract a few more times, I agree with Henry. I commented out the things that seemed to me superfluous. Regarding the last remark of "more gen and relations means more stability", I have counter examples of that. So I prefer to not state it, even heuristically, that way}We introduce a notion of \emph{efficient stability} for finite presentations of groups. Informally, a finite presentation using generators $S$ and relations $R$ is \emph{stable} if any map from $S$ to unitaries 
%(say, in some finite von Neumann algebra) 
that approximately satisfies the relations (in the tracial norm) is close to the restriction of a representation of $G$ to the subset $S$. This notion and variants thereof have been extensively studied in recent years, in part motivated by connections to property testing in computer science. The novelty in our work is the focus on \emph{efficiency}, which, informally, places an onus on small presentations --- in the sense of encoding length.  
%(of size, say, logarithmic in the group size)
The goal in this setup is to achieve non-trivial tradeoffs between the presentation length and its modulus of stability.
%---as, intuitively, more generators and relations place more constraints on the approximate homomorphism, making stability easier to achieve---whereas we wish to minimize the number of variables and constraints.

\mnote{Here I genuinely changed the phrasing. I don't think it is currently very good, but it is better than it was. Feel free to iterate:}\tnote{slightly edited}With this goal in mind we analyze various natural examples of presentations. We provide a general method for constructing presentations of $\Z_2^k$ from linear error-correcting codes. We observe that the resulting presentation has a weak form of stability exactly when the code is  \emph{locally testable}.\mnote{I have a problem with the notion of locally testable. The notion reffers ususally to testable LDPCs. We do not need the low density to define our notions and construction, only the testable part. I think we should change it to just testable codes... What do you think? This appears throughout the intro... It won't be a major change}\tnote{Right, it's not clear what motivation there is for the ``locally'' aspect in the language of groups; that the relations be ``short''---is that ever relevant? I'm ok for dropping ``locally''. It's not hard as we can search for the word.} \hnote{I'm fine with dropping ``locally''. The key thing that makes the problem nontrivial is that the \emph{number} of relations (i.e. complexity of presentation) is small, right?} This raises the question of whether locally testable codes give rise to genuinely stable presentations using this method.  While we cannot show that this is the case in general, we leverage recent results in the study of non-local games in quantum information theory (Ji et al., Discrete Analysis 2021) to show that a specific instantiation of our construction, based on the Reed-Muller family of codes, leads to a stable presentation of $\Z_2^k$ of size $\poly\log(k)$ only. As an application, we combine this result with recent work of de la Salle (arXiv:2204.07084) to re-derive  the quantum low-degree test of Natarajan and Vidick (IEEE FOCS'18), which is a key building block in the recent refutation of Connes' Embedding Problem via complexity theory (Ji et al., arXiv:2001.04383). 
\end{abstract}

@@

A finer property which concerns us here is the \emph{soundness} of the code, a parameter that is connected to the notion of \emph{local testability}. A code can be (non-uniquely) specified through a \emph{parity-check matrix} $h\in\F^{m\times n}$ as $\code = \ker h$. The rows of $h$ are thought of as constraints (``parity checks'') that specify $\code$ as a subspace of $\F^n$. A code is called locally testable \emph{with soundness $\rho$} if for every $x\in \F^n$, $\frac{1}{m}|hx|\geq \rho \, \frac{1}{n}\, d(x,\code)$, where $d(x,\code)$ denotes the minimum of $|x-c|$ over $c\in \code$.  Ideally one would like to design families of good codes such that in addition $\rho$ is bounded below by a constant independent of $n$.\footnote{Note that in principle a code can have a small minimal distance $d$, and still be locally testable with soundness $\rho>0$. So a family of codes can be ``locally testable'' without being ``good.''} This is a challenging task, and the construction of families of good locally testable codes was only achieved very recently~\cite{LTC_DELLM,LTC_Panteleev_Kalachev}.

The terminology ``locally testable'' comes from an interpretation of $\frac{1}{m}|hx|$ as the probability of rejection of a natural ``tester'' for $\code$, i.e.\ an algorithm that on input $x$ checks a randomly chosen row $h_i$ of $h$ and accepts if and only if $h_i\cdot x = \sum_j h_{ij} x_j =0$.  Thus a code is called locally testable if words that are far from the code have a high probability of being rejected according to this tester. This notion plays a central role in applications of codes to complexity theory~\cite{babai1991non,PCP_thm}, and continues to be actively studied. See e.g.~\cite{LTC_DELLM,LTC_Panteleev_Kalachev} for a recent breakthrough on the topic.

We make an observation that connects the study of locally testable codes to questions of stability in group theory and motivates our work. Let $\code=\ker h$ be a linear code as above, and suppose for simplicity that $\F=\F_2$ is the binary field. Consider the finitely presented group 
\begin{align}
 G(h) \,=\, \langle S:R\rangle \,=\, \big\langle &x_1,\ldots,x_n \,:\quad  x_j^2=e\quad \forall 1\leq j\leq n\;,\notag\\
& \quad \prod_{1\leq j \leq n} x_j^{h_{ij}} = e\;,\quad [x_{j_1},x_{j_2}]^{h_{ij_1} h_{ij_2}}=e\quad \forall 1\leq i\leq m\,\forall 1\leq j_1,j_2\leq n\,\big\rangle\;.\label{eq:gh-intro}
\end{align}
Here a commutation relation between two generators  $[x_{j_1},x_{j_2}]=e$ is imposed only when needed for the relations $\prod_{1\leq j \leq n} x_j^{h_{ij}} = e$ to make sense, i.e.\ two generators are required to commute only if they both take part in the same equation, which is the case if and only if $h_{ij_1}h_{ij_2}=1$ for some $i$. \tnote{added:}While one could add all pairwise commutation relations, forcing $G(h)$ to be abelian --- and indeed we will do this in some cases later --- we choose the specific presentation~\eqref{eq:gh-intro} for consistency with the literature on non-local games, and particularly so-called \emph{linear constraint system games}, which we review in more detail below.

We observe that $1$-dimensional representations, i.e.\ maps from $S=\{x_1,\ldots,x_n\}$ to $\{-1,1\}$ that satisfy all relations $R$, are in one-to-one correspondence with elements of $\code$. Moreover, \emph{approximate} $1$-dimensional representations, i.e.\ maps from  $S=\{x_1,\ldots,x_n\}$ to $\{-1,1\}$ that satisfy a fraction $1-\eps$ of the matrix relations for small $\eps$, can be identified with words $x\in\F_2^n$ such that $\frac{1}{m}|hx|\leq \eps$.\footnote{We conditioned only on the matrix relations since the commutation and incolution relations are automatically satisfied by out choice of range $\{\pm1\}$. In principle we could allow the map to range over $U(\C)$ instead of $\{-1,1\}$, and in general we will allow this. But, for the purposes of this introduction, it is simpler to restrict to the $\{-1,1\}$-valued case.} In particular, we notice that $\code$ is locally testable with soundness $\rho$ if and only if $\eps$-approximate $1$-dimensional representations of $G(h)$ are $\nicefrac{C\eps}{\rho}$-close to genuine $1$-dimensional representations, where $C$ depends on the distribution we choose over the relations. E.g., if we check an involution relation with probability $\nicefrac{1}{3}$, a commutation relation with probability $\nicefrac{1}{3}$, and a matrix relation with probability $\nicefrac{1}{3}$, then $C=3$.

@@

We can now ask the question: is $G(h)$, the group presentation defined in~\eqref{eq:gh-intro}, stable according to Definition~\ref{def:eff-stab-intro}? 
One can verify that with $\mu_R$ which chooses with probability $\nicefrac{1}{3}$ whether to check an involution, a commutation or a matrix row,  and $\mu_S$ the uniform distribution, $G(h)$ is $(\delta,\mu_S,\mu_R,1)$-stable for $\delta=3\eps/\rho$ if and only if $\code = \ker h$ is locally testable with soundness $\rho>0$. But what about higher-dimensional approximate representations? Are these also stable, or does one need to make further requirements on $\code$ beyond local testability? We do not yet have a comprehensive answer to these questions. However, the answer cannot be straightforward: even deducing basic properties about the group $G(h)$ given its presentation -- let alone its stability properties -- appears to be a challenging task. For example, there are examples of parity check matrices $h$ for which the group $G(h)$ is non-abelian or even non-amenable; see Remark~\ref{rk:non-abelian}. In fact, every finitely generated group can be embedded into $G(h)$ for some $h$ (see \cite{slofstra2019set}).

@@

\tnote{modified from here:}
To simplify the problem let us consider the following presentation for an obviously finite and abelian group
\begin{align}
  \widetilde{G(h)} \,=\, \langle S:R\rangle \,=\, \big\langle &x_1,\ldots,x_n \,:\quad  x_j^2=e\quad \forall 1\leq j\leq n\;,\notag\\
 & \quad \prod_{1\leq j \leq n} x_j^{h_{ij}} = e\;,\quad [x_{j_1},x_{j_2}]=e\quad \forall 1\leq i\leq m\,\forall 1\leq j_1,j_2\leq n\,\big\rangle\;.\label{eq:wtgh-intro}
 \end{align}
This is the same as $G(h)$, except that all pairwise commutations have been added. As we show formally later (see Lemma~\ref{lem:com-code}), it is not hard to check that $\widetilde{G(h)}$ is a presentation of the group $\Z_2^k$, for $k=\dim\ker h$. Our most important results pertain to the stability of the presentation $\widetilde{G(h)}$; although we will later (Section~\ref{sec:braiding}) consider some non-abelian extensions of it.

@@

\mnote{The following two paragraphs are problematic. The size of $G(h)$ is independent of $n,k$ in general, even if it is finite. The only case in which you can bound the size of $G(h)$ according to $n,k$ is when it is \textbf{abelian}, and we insisted on not assuming that. I feel like we should add all commutation relations before discussing this part.}\tnote{Yes so I tried doing this above, is it better? If we keep the notation, we need to modify around Lemma~\ref{lem:com-code} as well}

@@

Can such ``efficient'' presentations of $G(h)$ still be stable? The naive approach, of extending an $\eps$-approximate homomorphism of $G(h)=\langle S:R\rangle$ into an $\eps'$-approximate homomorphism of the multiplication table presentation of $\Z_2^k$, leads to $\eps'=\Omega(k^2\eps)$ and hence a logarithmic dependence of the modulus of stability on the group size. Is it possible to do better? Glebsky's result for the case of $\Z^2$, which is an infinite group, suggests that in some cases the modulus of stability can be independent of the group size. It is therefore not clear what if any of the known group parameters should play a role in it in general.  
 \mnote{Actually, we have a lower bound on the stability rate of $\Z^d$. It is $\Omega(\eps^{1/d})$ by a work of Becker and Mosheiff (they look at the permutation setup, but their example works for unitaries). Thus, there is not better solution in general, without assuming that all the generators are involutions. I just sent to John this draft, and by how he was confused by this paragraph, and the fact the the last question is known to be flase, we should rewrite it.}\tnote{Ok, I edited a bit}

@@

We do not know of any other presentation of $\Z_2^k$, arguably one of the simplest groups one could think of, that is stable with similar parameters as the ones stated in the theorem. It would be very interesting to discover different such presentations, built from locally testable error-correcting codes or not, for this group or others.

@@

Now we see that to each matrix $h$ we have associated a group $G(h)$ \hnote{do we have to use the tilde version now?}, and a game $\game_h$. Moreover, and quite interestingly, there is a one-to-one correspondence between representations of $G(h)$ and \emph{perfect strategies} in $\game_h$, i.e.\ strategies that have success probability $1$ in the game.\footnote{This correspondence was established for finite-dimensional strategies, and finite-dimensional representations, in~\cite{cleve2014characterization}. Extensions to infinite-dimensional strategies and representations have appeared in~\cite{cleve2017perfect}. See also e.g.~\cite{kim2018synchronous} for generalizations to the broader class of \emph{synchronous games}.}
This correspondence enables one to ``embed'' group representations into quantum strategies, thereby forcing them to demonstrate a high level of complexity; this is the approach at the heart of~\cite{slofstra2019set}. Going further, for applications one is often required to understand not only optimal but also near-optimal strategies, whose success probability is e.g.\ $1-\eps$. The same correspondence associates to such strategies approximate homomorphisms of $G(h)$~\cite{slofstra2018entanglement}. Here once can see that measuring closeness on average over the choice of relation is a natural choice, which is all but forced by the definition of a game, where the questions are selected according to some pre-specified distribution.

To summarize, approximate stability results for $G(h)$ enable one to obtain structural results about near-optimal strategies in $\game_h$ --- such a result is known as a \emph{rigidity} result in quantum information. A rigidity result about a game called the quantum low-degree test~\cite{natarajan2018low} is at the heart of the proof of $\sf MIP^* = \sf RE$. The analysis of this test requires an efficient stability result for the Pauli group. Informally, the reason that the stability result needs to be for an efficient presentation of the Pauli group \mnote{We mix Pauli with Pauli all the time. In the main text we call them Pauli, and mention that they are a special case of a Heisenberg group. I like the Pauli name, since we ignore the y operators in this case, and I thought they are formally part of the Pauli group. But, since we already use notations with P throughout, it would be probably easier to change to Pauli here as well. }\tnote{Well the P is a macro, and sometimes we write ``for all $W\in\{X,Z\}$'', so if you think that WH would be more appropriate I think that we could do it}\mnote{I replaced here to Pauli, because the WH name appeared only locally here....}, as opposed to e.g.\ the multiplication table presentation, is because to obtain the final result it is necessary that the ``complexity'' of the test, or game, is smaller than the ``complexity'' of the object, or group being tested. Of course here we are loose about what we mean by ``complexity,'' and refer to~\cite{ji2021mip,vidickmip} for high-level explanations. The quantum low-degree test is described and analyzed in Section~\ref{sec:pbt}. We end by mentioning an open question: if one was able to obtain $|S_k|,|R_k|=\poly(k)$ while also having $\delta(\eps)=\poly(\eps)$ in the informal theorem stated above, then this result would likely, through the connection we just described, have  consequences for the efficient verification of quantum computations in the framework of interactive proof systems---see e.g.~\cite{coladangelo2019verifier,natarajan2023bounding} for a sample of known results in this direction.

@@

\paragraph{Open questions.}\tnote{added this paragraph}
We leave many questions open. A natural direction is to determine if there are simple sufficient conditions on the matrix $h$ that guarantee that the presentation $G(h)$ (\hnote{added:} or $\widetilde{G(h)}$) is stable. As we discussed, the condition that $\ker h$ is a testable code is equivalent to stability for $1$-dimensional permutations. It thus seems likely that testability is not a sufficient condition in general; can testability be determined by a combinatorial parameter of $h$? More generally, finding other examples of efficient stability seems of intrinsic interest, besides potential applications to property testing and the construction of nonlocal games. In a different direction, as mentioned in the previous paragraph an analogue of Theorem~\ref{thm:main-inf} for the case of representations in permutations could have important implications towards showing the existence of non-sofic groups~\cite{CL_part1,CL_part2,BCLV_subgroup_tests}\tnote{Michael, can you add a reference to the relevant paper of yours/ours?}\mnote{Done. But, it seems as if the second paragraph of the 'related works on stability' subsection already hints towards this open problem.}\tnote{Ok thanks, I reformulated. We could also remove the refs here and only mention them in the previous paragraph. I think it's ok but if you prefer that way you can edit.}

@@

We consider the following presentation for the Pauli group \mnote{This is not the correct presentation. We use only $(a,0)$ and $(0,b)$ as generators (and not all pairs as currently written). Also, the anti commutation relation is in the other direction isn't it? I don't want to mess something because I am not sure which presentation you were aiming for, but this has typos in it for sure. Lastly, the commutation with $J$ should be of pairs $(a,0)$ and $(0,b)$ and not just $a$ and $b$}\tnote{sorry, fixed it}:
\begin{align}
 \pauli_k \,=\, \big\langle \{J\}\cup  \{(a,0),(0,b): a,b\in \Z_2^k\} &: (a,0)^2=(0,b)^2=J^2=e,\; [(a,0),J]=[(0,b),J]=e,\notag\\
 &\quad(a,0)(a',0)=(a+a',0),(0,b)(0,b')=(0,b+b') \notag\\
&\quad  (a,0)(0,b) = J^{a\cdot b} (0,b)(a,0) \; \quad \forall a,b,a',b'\in \Z_2^k\big\rangle\;.\label{eq:almost_mult_table_pres_of_Pauli}
\end{align}
This presentation is not quite the multiplication table presentation (it has $2^{k+1}+1$ generators, whereas $|\pauli_k|=2^{2k+1}$), but it is not far from it.  
Applying Theorem~\ref{thm:gh} we obtain the following consequence.%, which we again state only for the case of a subclass of approximate homomorphisms and the class of finite-dimensional matrix algebras\tnote{again, not sure what we want ultimately}.

@@

\label{sec:pres-code}\mnote{Why the same headline for the subsection as the section itself? Maybe we should change}\tnote{Sure, I changed}

@@

For the remainder of this section we specialize the discussion to the case where $q=2$. 
We make the simple but key observation that a parity check matrix for a code of dimension $k$ implies a finite group presentation in the following way. Introduce $n$ generators $S=\{x_1,\ldots,x_n\}$. Each of the generators is required to be an involution: $x_i^2=e$. For each row $i\in \{1,\ldots,m\}$ of the parity check matrix $h$, introduce a relation 
\[ R_i\,:\; \prod_{1\leq j \leq n} x_j^{h_{ij}}=e\;, \]
that ``verifies'' the parity check associated with the $i$-th row of $h$. Finally, to guarantee that $R_i$ is independent of the order in which the $x_j$ are multiplied, whenever $j\neq j'$ are such that $h_{ij}$ and $h_{ij'}$ are both nonzero we require that $x_j$ and $x_{j'}$ commute. This can be written succinctly using a relation 
\[ R'_{ijj'}\,:\; [x_j,x_{j'}]^{h_{ij} h_{ij'}}=e\;.\]
The presentation obtained in this way defines a group $G=G(h)$, already introduced in~\eqref{eq:gh-intro} and which with the present notation reads
\begin{equation}\label{eq:def-gh-pres}
 G(h) \,=\, \big\langle x_1,\ldots,x_n \,:\, x_j^2=e\,,\; R_i\,,\; R'_{ijj'}\,,\quad \forall 1\leq i\leq m,\, 1\leq j< j' \leq n\big\rangle\;.
\end{equation}
Note that we made the dependence of $G(h)$ on $h$ explicit. This is because in general, the group defined in this way may depend on $h$, and not only on $\code$. If however we further impose \emph{all} pairwise commutation relations then we obtain the following.

@@

We end this section with an example, the \emph{Hadamard code}. This code can be defined for any  $t\geq 1$ and it is a $[T,t,T/2]_2$ linear code, where $T=2^t$. For simplicity we write  $\code_\had$ to denote this code, omitting $t$. The Hadamard code is the subspace of linear functionals from $\field^t$ to $\field$ out of all such functions. As a linear space, $\code_\had$ can be described as  $(a\cdot b)_{a\in \field^t} \in \field^{\field^t}\cong \field^T$, for all $b\in \field^t$, where $ a\cdot b=\sum_{i=1}^t a_ib_i$ is again the dot product modulo $2$.  \mnote{The last definition was more confusing than clarifying. I changed it. }

@@

Rubinfeld and Sudan~\cite{rubinfeld1996robust} (building on the work of Babai, Fortnow, and Lund~\cite{babai1991non}) showed that $h_{\RM}$ is indeed a parity check matrix for $\code_{\RM}$. Furthermore, they showed that the parity check matrix gives rise to a $(d+2)$-local tester for $\code_{\RM}$, whose soundness $\rho$ (as defined in the introduction) is at least $\frac{1}{6}$. \mnote{Clearly it is a tester, but shouldn't we remark the soundness parameter Rubinfeld--Sudan calculated?}\tnote{Sure, added}

>>>>

\begin{abstract}
We introduce a notion of \emph{efficient stability} for finite presentations of groups. Informally, a finite presentation using generators $S$ and relations $R$ is \emph{stable} if any map from $S$ to unitaries 
%(say, in some finite von Neumann algebra) 
that approximately satisfies the relations (in the tracial norm) is close to the restriction of a representation of $G$ to the subset $S$. This notion and variants thereof have been extensively studied in recent years, in part motivated by connections to property testing in computer science. The novelty in our work is the focus on \emph{efficiency}, which, informally, places an onus on small presentations --- in the sense of encoding length.  
%(of size, say, logarithmic in the group size)
The goal in this setup is to achieve non-trivial tradeoffs between the presentation length and its modulus of stability.
%---as, intuitively, more generators and relations place more constraints on the approximate homomorphism, making stability easier to achieve---whereas we wish to minimize the number of variables and constraints.

With this goal in mind we analyze various natural examples of presentations. We provide a general method for constructing presentations of $\Z_2^k$ from linear error-correcting codes. We observe that the resulting presentation has a weak form of stability exactly when the code is  \emph{testable}. This raises the question of whether testable codes give rise to genuinely stable presentations using this method.  While we cannot show that this is the case in general, we leverage recent results in the study of non-local games in quantum information theory (Ji et al., Discrete Analysis 2021) to show that a specific instantiation of our construction, based on the Reed-Muller family of codes, leads to a stable presentation of $\Z_2^k$ of size $\poly\log(k)$ only. As an application, we combine this result with recent work of de la Salle (arXiv:2204.07084) to re-derive  the quantum low-degree test of Natarajan and Vidick (IEEE FOCS'18), which is a key building block in the recent refutation of Connes' Embedding Problem via complexity theory (Ji et al., arXiv:2001.04383). 
\end{abstract}

@@

A finer property which concerns us here is the \emph{soundness} of the code, a parameter that is connected to the notion of \emph{testability}. A code can be (non-uniquely) specified through a \emph{parity-check matrix} $h\in\F^{m\times n}$ as $\code = \ker h$. The rows of $h$ are thought of as constraints (``parity checks'') that specify $\code$ as a subspace of $\F^n$. A code is called testable \emph{with soundness $\rho$} if for every $x\in \F^n$, $\frac{1}{m}|hx|\geq \rho \, \frac{1}{n}\, d(x,\code)$, where $d(x,\code)$ denotes the minimum of $|x-c|$ over $c\in \code$.\footnote{In the literature, the notion of \emph{local} testability is emphasized, where in addition the parity-check matrix is required to have rows of low Hamming weight. This requirement is less important for us, and so we de-emphasize it.}  Ideally one would like to design families of good codes such that in addition $\rho$ is bounded below by a constant independent of $n$.\footnote{Note that in principle a code can have a small minimal distance $d$, and still be testable with soundness $\rho>0$. So a family of codes can be ``testable'' without being ``good.''} This is a challenging task, and the construction of families of good (locally) testable codes was only achieved very recently~\cite{LTC_DELLM,LTC_Panteleev_Kalachev}.

The terminology ``testable'' comes from an interpretation of $\frac{1}{m}|hx|$ as the probability of rejection of a natural ``tester'' for $\code$, i.e.\ an algorithm that on input $x$ checks a randomly chosen row $h_i$ of $h$ and accepts if and only if $h_i\cdot x = \sum_j h_{ij} x_j =0$.  Thus a code is called testable if words that are far from the code have a high probability of being rejected according to this tester. This notion (when accompanied by the locality constraint) plays a central role in applications of codes to complexity theory~\cite{babai1991non,PCP_thm}, and continues to be actively studied. See e.g.~\cite{LTC_DELLM,LTC_Panteleev_Kalachev} for a recent breakthrough on the topic.

We make an observation that connects the study of testable codes to questions of stability in group theory and motivates our work. Let $\code=\ker h$ be a linear code as above, and suppose for simplicity that $\F=\F_2$ is the binary field. Consider the finitely presented group 
\begin{align}
 G(h) \,=\, \langle S:R\rangle \,=\, \big\langle &x_1,\ldots,x_n \,:\quad  x_j^2=e\quad \forall 1\leq j\leq n\;,\notag\\
& \quad \prod_{1\leq j \leq n} x_j^{h_{ij}} = e\;,\quad [x_{j_1},x_{j_2}]^{h_{ij_1} h_{ij_2}}=e\quad \forall 1\leq i\leq m\,\forall 1\leq j_1,j_2\leq n\,\big\rangle\;.\label{eq:gh-intro}
\end{align}
Here a commutation relation between two generators  $[x_{j_1},x_{j_2}]=e$ is imposed only when needed for the relations $\prod_{1\leq j \leq n} x_j^{h_{ij}} = e$ to make sense, i.e.\ two generators are required to commute only if they both take part in the same equation, which is the case if and only if $h_{ij_1}h_{ij_2}=1$ for some $i$. \tnote{added:}While one could add all pairwise commutation relations, forcing $G(h)$ to be abelian --- and indeed we will do this in some cases later --- we choose the specific presentation~\eqref{eq:gh-intro} for consistency with the literature on non-local games, and particularly so-called \emph{linear constraint system games}, which we review in more detail below.

We observe that $1$-dimensional representations, i.e.\ maps from $S=\{x_1,\ldots,x_n\}$ to $\{-1,1\}$ that satisfy all relations $R$, are in one-to-one correspondence with elements of $\code$. Moreover, \emph{approximate} $1$-dimensional representations, i.e.\ maps from  $S=\{x_1,\ldots,x_n\}$ to $\{-1,1\}$ that satisfy a fraction $1-\eps$ of the matrix relations for small $\eps$, can be identified with words $x\in\F_2^n$ such that $\frac{1}{m}|hx|\leq \eps$.\footnote{We conditioned only on the matrix relations since the commutation and incolution relations are automatically satisfied by out choice of range $\{\pm1\}$. In principle we could allow the map to range over $U(\C)$ instead of $\{-1,1\}$, and in general we will allow this. But, for the purposes of this introduction, it is simpler to restrict to the $\{-1,1\}$-valued case.} In particular, we notice that $\code$ is testable with soundness $\rho$ if and only if $\eps$-approximate $1$-dimensional representations of $G(h)$ are $\nicefrac{C\eps}{\rho}$-close to genuine $1$-dimensional representations, where $C$ depends on the distribution we choose over the relations. E.g., if we check an involution relation with probability $\nicefrac{1}{3}$, a commutation relation with probability $\nicefrac{1}{3}$, and a matrix relation with probability $\nicefrac{1}{3}$, then $C=3$.

@@

We can now ask the question: is $G(h)$, the group presentation defined in~\eqref{eq:gh-intro}, stable according to Definition~\ref{def:eff-stab-intro}? 
One can verify that with $\mu_R$ which chooses with probability $\nicefrac{1}{3}$ whether to check an involution, a commutation or a matrix row,  and $\mu_S$ the uniform distribution, $G(h)$ is $(\delta,\mu_S,\mu_R,1)$-stable for $\delta=3\eps/\rho$ if and only if $\code = \ker h$ is testable with soundness $\rho>0$. But what about higher-dimensional approximate representations? Are these also stable, or does one need to make further requirements on $\code$ beyond local testability? We do not yet have a comprehensive answer to these questions. However, the answer cannot be straightforward: even deducing basic properties about the group $G(h)$ given its presentation -- let alone its stability properties -- appears to be a challenging task. For example, there are examples of parity check matrices $h$ for which the group $G(h)$ is non-abelian or even non-amenable; see Remark~\ref{rk:non-abelian}. In fact, every finitely generated group can be embedded into $G(h)$ for some $h$ (see \cite{slofstra2019set}).

@@

To simplify the problem let us consider the following presentation for an obviously finite and abelian group
\begin{align}
  \widetilde{G(h)} \,=\, \langle S:R\rangle \,=\, \big\langle &x_1,\ldots,x_n \,:\quad  x_j^2=e\quad \forall 1\leq j\leq n\;,\notag\\
 & \quad \prod_{1\leq j \leq n} x_j^{h_{ij}} = e\;,\quad [x_{j_1},x_{j_2}]=e\quad \forall 1\leq i\leq m\,\forall 1\leq j_1,j_2\leq n\,\big\rangle\;.\label{eq:wtgh-intro}
 \end{align}
This is the same as $G(h)$, except that all pairwise commutations have been added. As we show formally later (see Lemma~\ref{lem:com-code}), it is not hard to check that $\widetilde{G(h)}$ is a presentation of the group $\Z_2^k$, for $k=\dim\ker h$. Our most important results pertain to the stability of the presentation $\widetilde{G(h)}$; although we will later (Section~\ref{sec:braiding}) consider some non-abelian extensions of it.

@@

@@

Can such ``efficient'' presentations of $G(h)$ still be stable? The naive approach, of extending an $\eps$-approximate homomorphism of $G(h)=\langle S:R\rangle$ into an $\eps'$-approximate homomorphism of the multiplication table presentation of $\Z_2^k$, leads to $\eps'=\Omega(k^2\eps)$ and hence a logarithmic dependence of the modulus of stability on the group size. Is it possible to do better? Glebsky's result for the case of $\Z^2$, which is an infinite group, suggests that in some cases the modulus of stability can be independent of the group size. It is therefore not clear what if any of the known group parameters should play a role in it in general.

@@

We do not know of any other presentation of $\Z_2^k$, arguably one of the simplest groups one could think of, that is stable with similar parameters as the ones stated in the theorem. It would be very interesting to discover different such presentations, built from testable error-correcting codes or not, for this group or others.

@@

Now we see that to each matrix $h$ we have associated a group $G(h)$, and a game $\game_h$.\footnote{(A similar correspondence holds between $\wtG$ and a natural associated game $\widetilde{\game}_h$; our main results apply to the latter, but the present discussion is more general and applies to both.} Moreover, and quite interestingly, there is a one-to-one correspondence between representations of $G(h)$ and \emph{perfect strategies} in $\game_h$, i.e.\ strategies that have success probability $1$ in the game.\footnote{This correspondence was established for finite-dimensional strategies, and finite-dimensional representations, in~\cite{cleve2014characterization}. Extensions to infinite-dimensional strategies and representations have appeared in~\cite{cleve2017perfect}. See also e.g.~\cite{kim2018synchronous} for generalizations to the broader class of \emph{synchronous games}.}
This correspondence enables one to ``embed'' group representations into quantum strategies, thereby forcing them to demonstrate a high level of complexity; this is the approach at the heart of~\cite{slofstra2019set}. Going further, for applications one is often required to understand not only optimal but also near-optimal strategies, whose success probability is e.g.\ $1-\eps$. The same correspondence associates to such strategies approximate homomorphisms of $G(h)$~\cite{slofstra2018entanglement}. Here once can see that measuring closeness on average over the choice of relation is a natural choice, which is all but forced by the definition of a game, where the questions are selected according to some pre-specified distribution.

To summarize, approximate stability results for $G(h)$ enable one to obtain structural results about near-optimal strategies in $\game_h$ --- such a result is known as a \emph{rigidity} result in quantum information. A rigidity result about a game called the quantum low-degree test~\cite{natarajan2018low} is at the heart of the proof of $\sf MIP^* = \sf RE$. The analysis of this test requires an efficient stability result for the Pauli group. Informally, the reason that the stability result needs to be for an efficient presentation of the Pauli group, as opposed to e.g.\ the multiplication table presentation, is because to obtain the final result it is necessary that the ``complexity'' of the test, or game, is smaller than the ``complexity'' of the object, or group being tested. Of course here we are loose about what we mean by ``complexity,'' and refer to~\cite{ji2021mip,vidickmip} for high-level explanations. The quantum low-degree test is described and analyzed in Section~\ref{sec:pbt}. We end by mentioning an open question: if one was able to obtain $|S_k|,|R_k|=\poly(k)$ while also having $\delta(\eps)=\poly(\eps)$ in the informal theorem stated above, then this result would likely, through the connection we just described, have  consequences for the efficient verification of quantum computations in the framework of interactive proof systems---see e.g.~\cite{coladangelo2019verifier,natarajan2023bounding} for a sample of known results in this direction.

@@

\paragraph{Open questions.}
We leave many questions open. A natural direction is to determine if there are simple sufficient conditions on the matrix $h$ that guarantee that the presentation $G(h)$ (or $\widetilde{G(h)}$) is stable. As we discussed, the condition that $\ker h$ is a testable code is equivalent to stability for $1$-dimensional permutations. It thus seems likely that testability is not a sufficient condition in general; can testability be determined by a combinatorial parameter of $h$? More generally, finding other examples of efficient stability seems of intrinsic interest, besides potential applications to property testing and the construction of nonlocal games. In a different direction, as mentioned in the previous paragraph an analogue of Theorem~\ref{thm:main-inf} for the case of representations in permutations could have important implications towards showing the existence of non-sofic groups~\cite{CL_part1,CL_part2,BCLV_subgroup_tests}.

@@

We consider the following presentation for the Pauli group:
\begin{align}
 \pauli_k \,=\, \big\langle \{J\}\cup  \{(a,0),(0,b): a,b\in \Z_2^k\} &: (a,0)^2=(0,b)^2=J^2=e,\; [(a,0),J]=[(0,b),J]=e,\notag\\
 &\quad(a,0)(a',0)=(a+a',0),(0,b)(0,b')=(0,b+b') \notag\\
&\quad  (a,0)(0,b) = J^{a\cdot b} (0,b)(a,0) \; \quad \forall a,b,a',b'\in \Z_2^k\big\rangle\;.\label{eq:almost_mult_table_pres_of_Pauli}
\end{align}
This presentation is not quite the multiplication table presentation (it has $2^{k+1}+1$ generators, whereas $|\pauli_k|=2^{2k+1}$), but it is not far from it.  
Applying Theorem~\ref{thm:gh} we obtain the following consequence.%, which we again state only for the case of a subclass of approximate homomorphisms and the class of finite-dimensional matrix algebras\tnote{again, not sure what we want ultimately}.

@@

\label{sec:pres-code}

@@

For the remainder of this section we specialize the discussion to the case where $q=2$. 
We make the simple but key observation that a parity check matrix for a code of dimension $k$ implies a finite group presentation in the following way. Introduce $n$ generators $S=\{x_1,\ldots,x_n\}$. Each of the generators is required to be an involution: $x_i^2=e$. For each row $i\in \{1,\ldots,m\}$ of the parity check matrix $h$, introduce a relation 
\[ R_i\,:\; \prod_{1\leq j \leq n} x_j^{h_{ij}}=e\;, \]
that ``verifies'' the parity check associated with the $i$-th row of $h$. Finally, to guarantee that $R_i$ is independent of the order in which the $x_j$ are multiplied, whenever $j\neq j'$ are such that $h_{ij}$ and $h_{ij'}$ are both nonzero we require that $x_j$ and $x_{j'}$ commute. This can be written succinctly using a relation 
\[ R'_{ijj'}\,:\; [x_j,x_{j'}]^{h_{ij} h_{ij'}}=e\;.\]
The presentation obtained in this way defines a group $G=G(h)$, already introduced in~\eqref{eq:gh-intro} and which with the present notation reads
\begin{equation}\label{eq:def-gh-pres}
 G(h) \,=\, \big\langle x_1,\ldots,x_n \,:\, x_j^2=e\,,\; R_i\,,\; R'_{ijj'}\,,\quad \forall 1\leq i\leq m,\, 1\leq j< j' \leq n\big\rangle\;.
\end{equation}
Note that we made the dependence of $G(h)$ on $h$ explicit. This is because in general, the group defined in this way may depend on $h$, and not only on $\code$. If however we further impose \emph{all} pairwise commutation relations, as in~\eqref{eq:wtgh-intro}, then we obtain the following.

@@

We end this section with an example, the \emph{Hadamard code}. This code can be defined for any  $t\geq 1$ and it is a $[T,t,T/2]_2$ linear code, where $T=2^t$. For simplicity we write  $\code_\had$ to denote this code, omitting $t$. The Hadamard code is the subspace of linear functionals from $\field^t$ to $\field$ out of all such functions. As a linear space, $\code_\had$ can be described as  $(a\cdot b)_{a\in \field^t} \in \field^{\field^t}\cong \field^T$, for all $b\in \field^t$, where $ a\cdot b=\sum_{i=1}^t a_ib_i$ is again the dot product modulo $2$.

@@

Rubinfeld and Sudan~\cite{rubinfeld1996robust} (building on the work of Babai, Fortnow, and Lund~\cite{babai1991non}) showed that $h_{\RM}$ is indeed a parity check matrix for $\code_{\RM}$. Furthermore, they showed that the parity check matrix gives rise to a $(d+2)$-local tester for $\code_{\RM}$, whose soundness $\rho$ (as defined in the introduction) is at least $\frac{1}{6}$.
